### 2012-10-26 inst/examples/testing-gp-sdp.Rmd
exploring hyperparameter estimation, etc

Quick trial SDP approaches with GP function
========================================================

```{r libraries}
require(pdgControl)
require(ggplot2)
opts_knit$set(upload.fun = socialR::flickr.url)
```

@@

```{r}
for(t in 1:(T-1))
  x[t+1] = z_g(sigma_g) * f(x[t], h=0, p=p)
```

@@

## Stochastic Dynamic programming solution based on the posterior Gaussian process:

Define a transition matrix $F$ from the Gaussian process, giving the probability of going from state $x_t$ to $x_{t+1}$.
We already have the Gaussian process mean and variance predicted for each point $x$ on our grid, so this is simply:

```{r}
V <- sqrt(diag(gp$Cf))
F <- sapply(x_grid, function(x) dnorm(x, gp$Ef, V))
F <- t(apply(F,1,function(x) x/sum(x))) # normalize
h_grid <- x_grid
```

True $f(x)$

```{r}
mu <- sapply(x_grid, f, 0, p)
F_true <- sapply(x_grid, function(x) dnorm(x, mu, sigma_g))
F_true <- t(apply(F_true,1,function(x) x/sum(x))) # normalize
```

Somewhat silly way to adjust the matrix by harvest level:
```{r}
n <- length(h_grid)
per_harvest <- function(F,n)
  lapply(0:(n-1), function(i){
  d <- diag(1, nrow=(n-i))
  top <- matrix(0, nrow=i, ncol=n)
  top[,1] <- 1
  side <- matrix(0, nrow=(n-i), ncol=i)
  out <- F %*% rbind(top, cbind(d,side))
  t(apply(out,1,function(x) x/sum(x))) # normalize
})

```

True F is okay but not quite right:

```{r}
matrices <- per_harvest(F_true,n)
opt <- find_dp_optim(matrices, x_grid, h_grid, 20, 0, profit, delta=.01)
plot(opt$D[,1])
```

whoops: gp inferred F isn't working:

```{r}
matrices <- per_harvest(F,n)
opt <- find_dp_optim(matrices, x_grid, h_grid, 20, 0, profit, delta=.01)
plot(opt$D[,1])
```

@@

```{r}
profit <- profit_harvest(price=1, c0 = 0, c1=0)
pdfn <- function(P, s) dlnorm(P, 0, s)
matrices <- determine_SDP_matrix(f, p, x_grid, h_grid, sigma_g, pdfn)
opt <- find_dp_optim(matrices, x_grid=x_grid, h_grid=h_grid, OptTime=20, xT=0, profit=profit, delta=.05)
plot(opt$D[,1])
```

>>>>

Quick trial SDP approaches with GP function
==============================================

```{r libraries}
require(pdgControl)
require(nonparametricbayes)
require(ggplot2)
opts_knit$set(upload.fun = socialR::flickr.url)
```

@@

```{r}
for(t in 1:(T-1))
  x[t+1] = z_g() * f(x[t], h=0, p=p)
```

@@

## Optimization over hyperparameters:

```{r}
minusloglik <- function(par){
  gp <- gp_fit(obs, X, par)
  -gp$llik
}
par <- c(sigma_n=1, l=1)
minusloglik(par)
o <- optim(par, minusloglik)
o
hyperpars <- o$par
```

**Yikes.**

let's try 1-D optimization:

```{r}
minusloglik <- function(par){
  gp <- gp_fit(obs, X, c(sigma_n=par, l=1))
  -gp$llik
}
minusloglik(1)
o <- optimize(minusloglik, c(0,150))
o
hyperpars <- c(sigma_n=o$minimum, l=1)
```

**Yikes, that shouldn't happen either**

Plot the optimal hyperparameter solution:

```{r}
gp <- gp_fit(obs, X, hyperpars)
df <- data.frame(x=X, y=gp$Ef, ymin=(gp$Ef-2*sqrt(abs(diag(gp$Cf)))), ymax=(gp$Ef+2*sqrt(abs(diag(gp$Cf)))))
true <- data.frame(x=X, y=sapply(X,f, 0, p))
ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
  geom_line(data=true, aes(x,y), col='red', lty=2)
```

### Misc ways to calc harvest matrices

@@

```{r}
profit <- profit_harvest(price=1, c0 = 0, c1=0)
pdfn <- function(P, s) dlnorm(P, 0, s)
sdp_mat <- determine_SDP_matrix(f, p, x_grid, h_grid, sigma_g, pdfn)
opt <- find_dp_optim(sdp_mat, x_grid=x_grid, h_grid=h_grid, OptTime=20, xT=0, profit=profit, delta=.05)
plot(opt$D[,1])
```

Somewhat silly way to adjust the matrix by harvest level:

```{r}
n <- length(h_grid)
silly <- function(F,n)
  lapply(0:(n-1), function(i){
  d <- diag(1, nrow=(n-i))
  top <- matrix(0, nrow=i, ncol=n)
  top[,1] <- 1
  side <- matrix(0, nrow=(n-i), ncol=i)
  out <- F %*% rbind(top, cbind(d,side))
  t(apply(out,1,function(x) x/sum(x))) # normalize
})
```

binning function, not quite right:

```{r}
bin <- function(x, grid_centers){ 
  match <- which.min(abs(grid_centers - x)) 
  out <- integer(length(grid_centers))
  out[match] <- 1
  out
  }
per_harvest <- function(F,n){
  lapply(h_grid, function(h){
    t(sapply(x_grid, function(x) bin(x - h, x_grid)))
  })
}
```
